Stateless Model Checking with Data-Race Preemption Points

### 2015-11-14 pldi-16/di.tex
touch up di a bit

Motivated by state-of-the-art tools' inflexibility to change their preemption points,
Iterative Deepening searches among different combinations of such points during a test,
deciding on-the-fly whether to pursue each resulting state space or, or to defer it in favour of smaller ones.
Different state spaces are generated based on {\em subsets} of the preemption points which prior work would use; for example, ``preempt on all calls to {\tt mutex\_lock} but not on {\tt mutex\_unlock}''.
We show a visualization in Figure~\ref{fig:id}.

@@

For the rest of the paper, we abbreviate {\em preemption point} (PP),
%For the remainder of the paper, we will abbreviate {\em preemption point} (PP),
{\em model checking} (MC),
{\em single-state-space model checking} (SSS-MC), % (i.e., the approach of prior work),
Dynamic Partial Order Reduction (DPOR), and {\em state space estimate} (ETA).

@@

With Iterative Deepening, this is a simple matter of creating a new state space with a PP enabled on the racing instructions by each thread, as shown in Algorithm~\ref{alg:handledatarace}.
Note that even though a data race may involve two different instructions, we add new state spaces with only one new PP at a time.
Rather than adding a single state space configured to preempt on both involved instructions,
i.e., $AB =$ \GetPPSet{$j_0$} $\cup$ $\alpha$ $\cup$ $\beta$,
we prefer to add multiple smaller jobs which have a higher chance of completing in time. i.e.,
$A =$ \GetPPSet{$j_0$} $\cup$ $\alpha$ and
$B =$ \GetPPSet{$j_0$} $\cup$ $\beta$.
If $A$ and $B$ bug-free, they will in turn add $AB$ later.
%We take care to avoid duplicating any superset state spaces with PPs on multiple data races.
The condition on line 1 ensures that we avoid duplicating any state spaces with multiple data-race PPs;
for example, $AB$ is reachable by multiple paths through its different subsets, but should be added only once.

@@

\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& \multicolumn{2}{c}{\texttt{struct x \{ int foo; int baz; \} *x;}} \\
	& \multicolumn{2}{c}{\texttt{struct y \{ int bar; \} *y;~~~~~~~~~~}} \\
	\\
	& Thread 1 & Thread 2 \\
	1 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	2 & \texttt{\hilight{olivegreen}{free}(x);} \\
	3 & & \texttt{// x's memory recycled} \\
	4 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	5 & & \texttt{// ...initialize...}\\
	6 & & \texttt{publish(y);} \\
	7 & & \texttt{\hilight{brickred}{y->bar = ...;}} \\
\end{tabular}
\caption{A common execution pattern with {\tt malloc()} that produces false positive data race candidates.}
\label{fig:recycle}
\end{figure}
\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& Thread 1 & Thread 2 \\
	1 & \texttt{publish(x);} & \\
	2 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	3 & \texttt{\hilight{olivegreen}{free}(x);} \\
	4 & & \texttt{x2 = get\_published\_x();} \\
	5 & & \texttt{// x's memory recycled} \\
	6 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	7 & & \texttt{\hilight{brickred}{x2->foo = ...;}} \\
\end{tabular}
\caption{If a single-pass data race detector discarded candidates matching the malloc-recycle pattern,
it would miss the bug in this adversarial program.}
\label{fig:recycle-bug}
\end{figure}

>>>>

Motivated by state-of-the-art tools' inflexibility to change their preemption points,
Iterative Deepening searches among different combinations of such points during a test,
deciding on-the-fly whether to pursue each resulting state space, or to defer it in favour of smaller ones.
Different state spaces are generated based on {\em subsets} of the preemption points which prior work would use; for example, ``preempt on all calls to {\tt mutex\_lock} but not on {\tt mutex\_unlock}''.
We show a visualization in Figure~\ref{fig:id}.

@@

For the rest of the paper, we will abbreviate {\em preemption point} (PP),
%For the remainder of the paper, we will abbreviate {\em preemption point} (PP),
{\em model checking} (MC),
{\em single-state-space model checking} (SSS-MC), % (i.e., the approach of prior work),
Dynamic Partial Order Reduction (DPOR), and {\em state space estimate} (ETA).

@@

With Iterative Deepening, this is a simple matter of creating a new state space with a PP enabled on the racing instructions by each thread, as shown in Algorithm~\ref{alg:handledatarace}.
Note that even though a data race may involve two different instructions, we add new state spaces with only one new PP at a time.
Rather than adding a single state space configured to preempt on both involved instructions,
i.e., $AB =$ \GetPPSet{$j_0$} $\cup$ $\alpha$ $\cup$ $\beta$,
we prefer to add multiple smaller jobs which have a higher chance of completing in time. i.e.,
$A =$ \GetPPSet{$j_0$} $\cup$ $\alpha$ and
$B =$ \GetPPSet{$j_0$} $\cup$ $\beta$.
If $A$ and $B$ are bug-free, they will in turn add $AB$ later.
%We take care to avoid duplicating any superset state spaces with PPs on multiple data races.
The condition on line 1 ensures that we avoid duplicating any state spaces with multiple data-race PPs;
for example, $AB$ is reachable by multiple paths through its different subsets, but should be added only once.

@@

\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& \multicolumn{2}{c}{\texttt{struct x \{ int foo; int baz; \} *x;}} \\
	& \multicolumn{2}{c}{\texttt{struct y \{ int bar; \} *y;~~~~~~~~~~}} \\
	\\
	& Thread 1 & Thread 2 \\
	1 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	2 & \texttt{\hilight{olivegreen}{free}(x);} \\
	3 & & \texttt{\hilight{commentblue}{// x's memory recycled}} \\
	4 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	5 & & \texttt{\hilight{commentblue}{// ...initialize...}}\\
	6 & & \texttt{publish(y);} \\
	7 & & \texttt{\hilight{brickred}{y->bar = ...;}} \\
\end{tabular}
\caption{A common execution pattern with {\tt malloc()} that produces false positive data race candidates.}
\label{fig:recycle}
\end{figure}
\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& Thread 1 & Thread 2 \\
	1 & \texttt{publish(x);} & \\
	2 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	3 & \texttt{\hilight{olivegreen}{free}(x);} \\
	4 & & \texttt{x2 = get\_published\_x();} \\
	5 & & \texttt{\hilight{commentblue}{// x's memory recycled}} \\
	6 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	7 & & \texttt{\hilight{brickred}{x2->foo = ...;}} \\
\end{tabular}
\caption{If a single-pass data race detector discarded candidates matching the malloc-recycle pattern,
it would miss the bug in this adversarial program.}
\label{fig:recycle-bug}
\end{figure}
